---
title: "Quantile Regression with `R`/`Stan`/`JAGS`"
output: html_document
---

```{r, echo=FALSE, message=FALSE, error=FALSE}
library(knitr)
knitr::opts_chunk$set(
  collapse = TRUE,
  cache = TRUE 
)
```


### Data
A quick data generating mechanism. Roughly speaking, the $95^{th}$ percentile of `beta` will be about 2.9, and will remain about 1 for `alpha`.

```{r}
# Set-up the data generating mechainsm
set.seed(666)
N     <- 500
x     <- runif(N, max=10)
alpha <- 1
beta  <- 2
y     <- alpha + beta * x + rnorm(N, sd = .6 * x)
p     <- 0.95

# The dataset to be used for estiamtion
data_set   <- list(y = y, x = x, p = p)
data_frame <- as.data.frame(data_set)[-3]

```


### Quantile Regression with the `quantreg` package in `R`
This fits pretty much instantaneously, and gets answers that look about right (alpha of about 1, beta of about 2.9).
```{r}
library(quantreg)
qreg_fit <- rq(y ~ x, data = data_frame, tau = p)
# (setting ci = TRUE seems to choke it for some reason)
 
# See results
summary(qreg_fit)
```


### Quantile Regression with the `bayesQR` package in `R`
This fits pretty much instantaneously, and seems to get sensible answers too.
```{r}
library(bayesQR)
sink(file="/dev/null") # Supress calls to cat(), for now
bqr_fit <- bayesQR(y ~ x, data = data_frame, quantile = p, ndraw = 10000)
sink()

# See results
(bqr_summary <- summary(bqr_fit))
```


### Quantile Regression in `JAGS`

```{r}
# Adapted from http://stats.stackexchange.com/q/17672
library(rjags)

jags_code <- "
  model{
    for(i in 1:length(y)){
      mu[i] <- alpha + beta * x[i]
      w[i]  ~ dexp(tau)
      me[i] <- (1 - 2 * p) / (p * (1 - p)) * w[i] + mu[i]
      pe[i] <- (p * (1 - p) * tau) / (2 * w[i])
      y[i]  ~ dnorm(me[i], pe[i])
    }
    
    # Regression Priors
    alpha ~ dnorm(0, 1E-6)
    beta  ~ dnorm(0, 1E-6)
    
    lsigma ~ dunif(-5, 15)
    sigma  <- exp(lsigma / 2)
    tau    <- pow(sigma, -2)
  }
"

# Init the model
n_iter <- 10000
jags_model <- jags.model(file = textConnection(jags_code), data = data_set, 
                       n.chains = 4, n.adapt = n_iter / 2)

# Run some MCMC iterations
params <- c("alpha", "beta", "sigma")
jags_samples <- coda.samples(jags_model, params, n.iter = n_iter)

# Results
t(apply(
  data.frame(do.call(rbind, jags_samples)), 2, function(x)
    c(mean = mean(x), quantile(x, c(0.005, 0.25, 0.5, 0.75, 0.95)))
))

```


### Quantile Regression in `Stan`
The model below converged with 5k iterations in less than 30 seconds.

```{r}
library(rstan)
options(mc.cores = parallel::detectCores())

scode <- "
data {
  int<lower=0> N;
  real<lower=0, upper=1> p;
  vector[N] x;
  vector[N] y;
}
parameters {
  real alpha;
  real beta;
  real<lower=0> sigma;
  vector<lower=0>[N] w;
}
transformed parameters {
  real<lower=0> tau;
  vector[N] mu;
  vector[N] me;
  vector[N] pe;
  vector[N] pe2;

  tau <- pow(sigma, -2);
  mu  <- alpha + beta * x;
  me  <- (1 - 2 * p) / (p * (1 - p)) * w + mu;
  pe  <- (p * (1 - p) * tau) ./ (2 * w);
  
  for(n in 1:N)
    pe2[n] <- inv_sqrt(pe[n]);
}
model {
  # Priors
  alpha ~ normal(0, 2);
  beta  ~ normal(0, 2);
  sigma ~ cauchy(0, 2.5);

  # Data Augmentation
  w     ~ exponential(tau);
  
  # The model
  y     ~ normal(me, pe2);
}
"

init_fun <- function(){
  list(alpha = 1 + rnorm(1, 0, 0.2), beta = 2.88 + rnorm(1, 0, 0.2), 
       sigma = runif(1, 0, 5), w = rexp(N, runif(N, 0, 4)))
}

params <- c("alpha", "beta", "sigma")
stan_fit <- stan(model_code = scode, data = within(data_set, N <- N), 
                 iter = 5000, verbose = FALSE,
                 pars = params)

print(stan_fit)

```


# Session Info
```{r}
sessionInfo()
```
